---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# circMeta2

Circular RNA (circRNA) is one kind of single-stranded, covalently closed noncoding RNA, which exerts important biological functions by acting as transcriptional regulators, microRNA sponges and protein templates. One important analysis of circRNAs is performing differential expression analysis of circRNAs across different biological conditions. Considering the data characteristics of circRNAs such as low read counts, the DE analysis of circRNAs imposes a unique challenge. To address the challenge, we develop a novel computational pipeline named circMeta2 to perform differential expression analysis of alternative back-splicing circRNAs in clusters to improve the statistical power by leveraging the often-overlooked additive effects of individual circRNAs in the clusters identified by the alternative back-splicing events, which may help improve the statistical power and downstream biological findings.

# Installation

You can install circMeta2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fzhaouf/circMeta2")
```


# samll sample size example

two samples from human cerebellum and two samples from human frontal cortex are used for demontrating small sample size usage. circRNAs are called using 'CIRCexplorer2 and output files are stored in data folder.


```{r, warning=FALSE, message=TRUE}

library(circMeta2)
# read circexplorer2 or ciri2 output files using file path, specify number of sample for each 
# condition using conditions parameter and the default is 2 samples for each condition.
circexplorers=c(
  '~/Library/CloudStorage/OneDrive-UniversityofFlorida/circRNA_project/Rpackage/circRNA/cerebellum/CIRCexplorer/SRR3192427.txt',
  '~/Library/CloudStorage/OneDrive-UniversityofFlorida/circRNA_project/Rpackage/circRNA/cerebellum/CIRCexplorer/SRR3192428.txt',
  '~/Library/CloudStorage/OneDrive-UniversityofFlorida/circRNA_project/Rpackage/circRNA/frontalcortex/CIRCexplorer/SRR3192424.txt',
  '~/Library/CloudStorage/OneDrive-UniversityofFlorida/circRNA_project/Rpackage/circRNA/frontalcortex/CIRCexplorer/SRR3192425.txt'
)

circ.obj = makecircObj(samplefiles = circexplorers, conditions = c(2,2), circ.method=c('CIRCexplorer2'),species='hg19')

#  clustering A5BS and A3BS circ-clusters
circ.obj = getCircCluster(circObj = circ.obj)

# individual circRNA DE using either pois Z for small sample or GLM for large sample with covariates
circ.obj = circRNADE(circObj = circ.obj, DEmethod = 'Pois')

# circ-cluster DE
results = circClusterDE(circObj=circ.obj, circ.cutoff=2, DEmethod='Meta')
head(results$A5BS.cluster)
head(results$A3BS.cluster)
```

# large sample size example

AD circRNA data from GRanges. 


```{r, warning=FALSE, message=TRUE}
data("BM10.circs", package = "circMeta2")
data("metainfo", package = "circMeta2")

# read circexplorer2 or ciri2 output files using file path, specify number of sample for each 
# condition using conditions parameter and the default is 2 samples for each condtion.
circ.obj = makecircObjfromGRanges(GRanges = BM10.circs,species='hg19',metadata=metainfo)

#  clustering A5BS and A3BS circ-clusters
circ.obj = getCircCluster(circObj = circ.obj)

# call inidivual circRNA DE 
circ.obj = circRNADE(circObj = circ.obj, DEmethod = 'GLM', formula_str = "readNumber ~ condid + age + sex")

# circ-cluster DE
results = circClusterDE(circObj=circ.obj, circ.cutoff=2, DEmethod='Meta')
head(results$A5BS.cluster)
head(results$A3BS.cluster)
```

